---
title: Making a map of African Swine Fever (ASF) outbreaks in Europe from the data
  scrapped from the OIE WAHID website
author: "Pavel Vodrážka"
date: "Thursday, January 15, 2015"
output:
  html_document: default
---

### Reason

The maps that can be obtained from the [OIE WAHID website](http://www.oie.int/wahis_2/public/wahid.php/Diseaseinformation/Diseaseoutbreakmaps) are not aesthetically 
pleasing, and can be downloaded only as raster images that are not suitable for 
including in print quality publications. Other sources of data and maps of ASF 
outbreaks were not found.

### Obtaining the data

There is no option to dowload the data or no public data API on the OIE WAHID 
website. Data needs to be scraped from the ["Immediate notifications and 
Follow-Ups"](http://www.oie.int/wahis_2/public/wahid.php/Diseaseinformation/Immsummary) part of the site.

We go all the way from getting the webpages of summaries of immediate 
notifications and follow-ups for african swine fever for all the years (2005 -- 
2014), extracting the links of the full country reports, getting the full 
reports (immediate notifications), extracting the links of the follow-up 
reports, getting the follow-up reports, and getting and processing data from the 
dowloaded reports.

The starting page is http://www.oie.int/wahis_2/public/wahid.php/Diseaseinformation/Immsummary. The following code refers to the version of the web denoted in the 
footer:
```
+--------------------------------------------------------------+
| World Animal Health Information Database (WAHID) – Version 1 |
| Copyright © World Organisation for Animal Health (OIE)       |
| Release date: 30 December 2013 August 2012                   |
+--------------------------------------------------------------+
```
```{r loadPackages}
library(rvest)
library(httr)
library(dplyr)
library(magrittr)
library(digest)
library(rworldmap)
library(lubridate)
```

```{r function-checkIfwebNotChanged, child="checkIfwebNotChanged.R", cache=TRUE}```

We first need to be able to programmatically fill the form on the page and download 
the resulting summaries. We prepare a function `getDiseaseSummaries` using some 
functions from the "RHTMLForms" package to accomplish this part of the task.

There are four forms on the page; after inspecting the page source we know that 
we are interested in the `diseaseform`. Our function `getDiseaseSummaries` 
contains a function `summaries_fun` that is created programmatically from the 
page source code and that will serve to submit that form and retrieve the 
resulting webpages.

```{r function-getDiseaseFormValues, child="getDiseaseFormValues.R", cache=TRUE, autodep=TRUE}```

```{r obtainFormValues, cache=TRUE, autodep=TRUE}
form_values <- getDiseaseFormValues(disease = "african swine fever")
```

```{r function-downloadSINF, child="downloadSINF.R", cache=TRUE, autodep=TRUE}```

```{r function-parseSINF, child="parseSINF.R", cache=TRUE, autodep=TRUE}```

```{r function-getCountries, child="getCountries.R", cache=TRUE, autodep=TRUE}```

```{r function-getSINF, child = "getSINF.R", cache=TRUE, autodep=TRUE}```

By inspecting the page source and the forms we figure out the necessary 
parameters to the `getDiseaseSummaries` function: `disease_id_terrestrial` 
(`= "12"` for ASF), and `year`. Now we obtain the yearly summaries of immediate 
notifications and follow-ups as a `summaries` object -- a list containing the raw 
html of pages returned by repeatedly submitting the form with given disease selected 
for the range of years speficied.

```{r getDiseaseSummaries, cache=TRUE, autodep=TRUE}
summaries <- getSINF(formValues = form_values
#                      ,years = 2012:2015
                     #                                 ,countryNames = "Africa"
                     #                                 ,newDownload = TRUE
#                      , cacheInterval = "1 min"
)

write.csv2(summaries, file = "summaries.csv", row.names = FALSE, na = "", fileEncoding = "UTF-8")
```

Now we will parse the html of the pages stored in the `summaries` object.

[//]: # (Finish here)

```{r function-downloadSO, child="downloadSO.R", cache=TRUE, autodep=TRUE}```

```{r function-parseSO, child="parseSO.R", cache=TRUE, autodep=TRUE}```

```{r function-getSO, child="getSO.R", cache=TRUE, autodep=TRUE}```

```{r getSO, cache=TRUE, autodep=TRUE}
summaries_of_outbreaks <- getSO(summaries)
write.csv2(summaries_of_outbreaks, file = "summaries_of_outbreaks.csv",
            row.names = FALSE,
            na = "",
            fileEncoding = "UTF-8")
```

```{r function-downloadOutbreakReport, child="downloadOutbreakReport.R", cache=TRUE, autodep=TRUE}```

```{r function-parseOutbreakReport, child="parseOutbreakReport.R", cache=TRUE, autodep=TRUE}```

```{r function-getOutbreakReports, child="getOutbreakReports.R", cache=TRUE, autodep=TRUE}```

```{r getOutbreakReports, cache=TRUE, autodep=TRUE}
outbreak_reports <- getOutbreakReports(summaries_of_outbreaks)
write.csv2(outbreak_reports, file = "outbreak_reports.csv", row.names = FALSE,
           na = "", fileEncoding = "UTF-8")
```

```{r examplePlot, cache=TRUE, autodep=TRUE}
# library(plotGoogleMaps)
# outbreak_reportsSPDF <- outbreak_reports
# coordinates(outbreak_reportsSPDF) <- ~lon + lat
# proj4string(outbreak_reportsSPDF) <- CRS("+proj=longlat +datum=WGS84")
# ic <- iconlabels(outbreak_reportsSPDF$species, height = 5, icon = TRUE)
# map <- plotGoogleMaps(outbreak_reportsSPDF,
#                       file = "map.html",
#                       zcol = "start_date",
#                       mapTypeId = "TERRAIN",
#                       legend = FALSE,
#                       control = FALSE,
#                       iconMarker = ic)

# library(ggmap)
# theme_set(theme_bw(16))
# europe_map <- qmap("Ukraine", zoom = 5, source = "stamen")
# europe_map +
#         geom_point(aes(x = lon, y = lat, colour = start_date, size = cases_total),
#                    data = outbreak_reports)
# qmplot(lon, lat, data = outbreak_reports)

```

Download and process event summaries or full reports, if needed (contain extra information).
Cite packages!
Make a consistent flow!
Plots (animation).
Save data to a database. (SQLite)
rworldmap::countryRegions
spacetime
