---
title: Making a map of African Swine Fever (ASF) outbreaks in Europe from the data
  scrapped from the OIE WAHID website
author: "Pavel Vodrážka"
date: "Thursday, January 15, 2015"
output:
  html_document: default
---

### Reason

The maps that can be obtained from the [OIE WAHID website](http://www.oie.int/wahis_2/public/wahid.php/Diseaseinformation/Diseaseoutbreakmaps) are not aesthetically 
pleasing, and can be downloaded only as raster images that are not suitable for 
including in print quality publications. Other sources of data and maps of ASF 
outbreaks were not found.

### Obtaining the data

There is no option to dowload the data or no public data API on the OIE WAHID 
website. Data needs to be scraped from the ["Immediate notifications and 
Follow-Ups"](http://www.oie.int/wahis_2/public/wahid.php/Diseaseinformation/Immsummary) part of the site.

We go all the way from getting the webpages of summaries of immediate 
notifications and follow-ups for african swine fever for all the years (2005 -- 
2014), extracting the links of the full country reports, getting the full 
reports (immediate notifications), extracting the links of the follow-up 
reports, getting the follow-up reports, and getting and processing data from the 
dowloaded reports.

The starting page is http://www.oie.int/wahis_2/public/wahid.php/Diseaseinformation/Immsummary. The following code refers to the version of the web denoted in the 
footer:
```
+--------------------------------------------------------------+
| World Animal Health Information Database (WAHID) – Version 1 |
| Copyright © World Organisation for Animal Health (OIE)       |
| Release date: 30 December 2013 August 2012                   |
+--------------------------------------------------------------+
```
```{r loadPackages, cache=TRUE}
library(rvest)
library(httr)
library(dplyr)
library(magrittr)
library(digest)
```

```{r function-checkIfwebNotChanged, cache=TRUE, autodep=TRUE}
checkIfwebNotChanged <- function(hashFile = "hashes.RData") {
        url <- "http://www.oie.int/wahis_2/public/wahid.php/Diseaseinformation/Immsummary"
        val <- list()
        resp1 <- GET(url)
        stop_for_status(resp1)
        doc1 <- html(resp1, encoding = "UTF-8")
        val[["disease form attributes"]] <- doc1 %>%
                html_nodes("#diseaseform") %>%
                html_attrs %>%
                digest(algo = "md5")
        val[["diseaseform scripts"]] <- doc1 %>%
                html_nodes("#diseaseform script") %>%
                digest(algo = "md5")
        val[["diseaseform inputs"]] <- doc1 %>%
                html_nodes("#diseaseform input") %>%
                digest(algo = "md5")
        val[["diseaseform selects' attributes"]] <- doc1 %>%
                html_nodes("select") %>%
                html_attrs %>%
                digest(algo = "md5")
        resp2 <- GET(paste0(url, "/listoutbreak"))
        stop_for_status(resp2)
        doc2 <- html(resp2, encoding = "UTF-8")
        val[["outbreakreport form attributes"]] <- doc2 %>%
                html_nodes("form[name='outbreakreport']") %>%
                html_attrs %>%
                digest(algo = "md5")
        val[["outbreakreport form inputs"]] <- doc2 %>%
                html_nodes("form[name='outbreakreport'] input") %>%
                digest(algo = "md5")
        
        if(file.exists(hashFile)) {
                saved <- dget(hashFile)
                comparison <- mapply(identical, val, saved)
                if(!all(comparison)) {
                        not_equal <- names(comparison[!comparison])
                        warning(paste0("There are changes on the OIE WAHID website that can affect the functionality of this package.\nThings that have changed are:\n- ",
                                       paste0(not_equal, collapse = "\n- ")),
                                immediate. = TRUE)
                        web_not_changed <<- FALSE
                        return(FALSE)
                        } else {
                                web_not_changed <<- TRUE
                                return(TRUE)
                                }
                } else {
                        warning("Impossible to check whether the OIE WAHID website has not changed (hash file not present, creating a new one from the current state of the site).")
                        dput(val, file = "hashes.RData")
                        web_not_changed <<- TRUE
                        return(TRUE)
                        }
        }       
```
We first need to be able to programmatically fill the form on the page and download 
the resulting summaries. We prepare a function `getDiseaseSummaries` using some 
functions from the "RHTMLForms" package to accomplish this part of the task.

There are four forms on the page; after inspecting the page source we know that 
we are interested in the `diseaseform`. Our function `getDiseaseSummaries` 
contains a function `summaries_fun` that is created programmatically from the 
page source code and that will serve to submit that form and retrieve the 
resulting webpages.

```{r function-getDiseaseFormValues, autodep=TRUE}
getDiseaseFormValues <- function(disease = character(),
                                 printOnly = FALSE,
                                 setGlobalOnly = FALSE) {
        if(!exists("web_not_changed"))
                web_not_changed <- checkIfwebNotChanged()
        if(!web_not_changed)
                warning("The OIE WAHID website has changed, the following may not work.",
                        immediate. = TRUE)
        url <- "http://www.oie.int/wahis_2/public/wahid.php/Diseaseinformation/Immsummary"
        resp <- GET(url)
        stop_for_status(resp)
        doc <- html(resp, encoding = "UTF-8")
        years_available <<- doc %>%
                html_nodes("#diseaseform #year option") %>%
                html_attr("value")
        terrestrial_values <- doc %>%
                html_nodes("#diseaseform #disease_id_terrestrial option") %>%
                html_attr("value")
        terrestrial_labels <- doc %>%
                html_nodes("#diseaseform #disease_id_terrestrial option") %>%
                html_attr("label")
        aquatic_values <- doc %>%
                html_nodes("#diseaseform #disease_id_aquatic option") %>%
                html_attr("value")
        aquatic_labels <- doc %>%
                html_nodes("#diseaseform #disease_id_aquatic option") %>%
                html_attr("label")
        values_labels <<- data.frame(label = c(terrestrial_labels, aquatic_labels),
                                     disease_id_hidden = c(terrestrial_values,
                                                           aquatic_values),
                                     disease_type_hidden = c(rep.int("0",
                                                                     length(terrestrial_values)),
                                                             rep.int("1",
                                                                     length(aquatic_values))
                                                             ),
                                     stringsAsFactors = FALSE
                                     ) %>%
                filter(disease_id_hidden != "-999") %>%
                arrange(label)
        class(values_labels) <- c(class(values_labels), "OIE diseaseform values_labels")
        if(setGlobalOnly)
                return(NULL)
        if(length(disease) > 0) {
                values_labels_filtered <- values_labels[grep(disease,
                                                             values_labels$label,
                                                             ignore.case = TRUE)
                                                        ,]
                }
        if(printOnly) {
                message("Values for submitting the \"diseaseform\" on the OIE WAHID 
                        \"Summary of Immediate notifications and Follow-ups\"
                        webpage [http://www.oie.int/wahis_2/public/wahid.php/Diseaseinformation/Immsummary)]\n")
                if(length(disease) > 0) {
                        message(paste0("Entered search for disease: \"", disease, "\"\n"))
                        if(nrow(values_labels_filtered) == 0) {
                                message("No disease found. Diseases available: \n")
                                message(paste0(names(values_labels), collapse = ", "))
                                message(paste(apply(values_labels, 1, paste, collapse = ", "),
                                              collapse = "\n"))
                                } else {
                                        message(paste0(names(values_labels_filtered), collapse = ", "))
                                        message(paste(apply(values_labels_filtered, 1, paste, collapse = ", "),
                                                      collapse = "\n"))
                                        }
                        } else {
                                message(paste0(names(values_labels), collapse = ", "))
                                message(paste(apply(values_labels, 1, paste, collapse = ", "),
                                              collapse = "\n"))
                                }
                message("\nYears available:\n")
                message(paste0(range(years_available), collapse = " - "))
                return(NULL)
                } else {
                        if(length(disease > 0)) {
                                if(nrow(values_labels_filtered) > 1)
                                        warning(paste0("Your search for pattern \"",
                                                       disease,
                                                       "\" returned more than one disease."),
                                                immediate. = TRUE)
                                if(nrow(values_labels_filtered) == 0) {
                                        warning(paste0("Your search for pattern\"",
                                                       disease,
                                                       "\" returned no disease."),
                                                immediate. = TRUE)
                                        return(NULL)
                                        }
                                return(values_labels_filtered)
                                } else {
                                        return(values_labels)
                                        }        
                        }
        }
```

```{r obtainFormValues, cache=TRUE, autodep=TRUE}
form_values <- getDiseaseFormValues(disease = "african swine fever")
```

```{r function-downloadSummaryOfImmediateNotificationsAndFollowUps, cache=TRUE, autodep=TRUE}
downloadSummaryOfImmediateNotificationsAndFollowUps <- function(diseaseID,
                                                                diseaseType,
                                                                year
                                                                ) {
        if(any(missing(diseaseID),
               missing(diseaseType),
               missing(year)))
                stop("Not all required arguments specified.")
        if(!exists("web_not_changed"))
                web_not_changed <- checkIfwebNotChanged()
        if(!web_not_changed)
                warning("The OIE WAHID website has changed, the following may not work.",
                        immediate. = TRUE)
        if(!exists("values_labels") | !exists("years_available"))
                getDiseaseFormValues(setGlobalOnly = TRUE)
        diseaseID_Type <- paste(diseaseID, diseaseType, sep = "-")
        diseaseIDs_Types <- apply(values_labels[, c("disease_id_hidden",
                                                    "disease_type_hidden")],
                                  1,
                                  paste0,
                                  collapse = "-")
        if(!diseaseID_Type %in% diseaseIDs_Types)
                stop(paste0("The entered combination of diseaseID and diseaseType (\"",
                            diseaseID_Type,
                            "\") doesn't exist."))
        if(!as.character(year) %in% years_available)
                stop(paste0("The year entered (\"",
                            year,
                            "\") is outside the range of WAHIS reports."))
        url <- "http://www.oie.int/wahis_2/public/wahid.php/Diseaseinformation/Immsummary"
        resp <- POST(url = url,
             body = paste0("disease_id_hidden=",
                           diseaseID,
                           "&disease_type_hidden=",
                           diseaseType,
                           "&year=",
                           year
                           ),
             content_type("application/x-www-form-urlencoded"))
        stop_for_status(resp)
        if(!exists("D1counter")) D1counter <<- 0
        D1counter <<- D1counter + 1
        if(D1counter %% 50 == 0) {
                cat(D1counter, "\n")
                } else {
                        if(D1counter %% 10 == 0) cat("|") else cat(".")
                }
        return(resp)
        }
```

```{r function-parseSummaryOfImmediateNotificationsAndFollowUps, cache=TRUE, autodep=TRUE}
parseSummaryOfImmediateNotificationsAndFollowUps <- function(response) {
        
        if(!"response" %in% class(response))
                stop ("The argument specified is not a response object.")
        url <- "http://www.oie.int/wahis_2/public/wahid.php/Diseaseinformation/Immsummary"
        postfields <- rawToChar(response[["request"]][["body"]][["config"]][["postfields"]])
        if(!all(response["url"] == url,       
                grep("disease_id_hidden=", postfields),
                grep("disease_type_hidden=", postfields),
                grep("year=", postfields)
                )) stop (paste0("The argument specified is not a summary of immediate notifications and follow-ups ", url))
        doc <- html(response, encoding = "UTF-8")
        country <- doc %>%
                html_nodes(".outbreakdetails .outbreak_country") %>%
                html_text(trim = TRUE) %>%
                gsub("[^A-Za-z ']", "", .)
        while(length((empty_countries <- which(nchar(country) == 0))) > 0) {
                country[empty_countries] <- country[empty_countries - 1]
                }
        status <- doc %>%
                html_nodes(".outbreakdetails :nth-child(2)") %>%
                html_text(trim = TRUE) %>%
                gsub("[0-9/]", "", .)
        date <- doc %>%
                html_nodes(".outbreakdetails :nth-child(2)") %>%
                html_text(trim = TRUE) %>%
                gsub("[A-Za-z ]", "", .) %>%
                as.Date("%d/%m/%Y")
        summary_country <- doc %>%
                html_nodes("#diseaseform :nth-child(3) a") %>%
                html_attr("onclick") %>%
                gsub("outbreaklist\\('|',[0-9]*);", "", .)
        reportid <- doc %>%
                html_nodes("#diseaseform :nth-child(3) a") %>%
                html_attr("onclick") %>%
                gsub("outbreaklist\\('[A-Z]*',|);", "", .)
        event_summary_link <- doc %>%
                html_nodes(".vacborder:nth-child(4) a") %>%
                html_attr("href") %>%
                paste0("http://www.oie.int", .)
        full_report_link <- doc %>%
                html_nodes(".vacborder:nth-child(5) a") %>%
                html_attr("href") %>%
                gsub("javascript: open_report\\(\"", "http://www.oie.int", .) %>%
                gsub("\",", "reportid=", .) %>%
                gsub("\\);", "", .)
        if(!exists("P1counter")) P1counter <<- 0
        P1counter <<- P1counter + 1
        if(P1counter %% 50 == 0) {
                cat(P1counter, "\n")
                } else {
                        if(P1counter %% 10 == 0) cat("|") else cat(".")
                }
        return(data.frame(country,
                          status,
                          date,
                          summary_country,
                          reportid,
                          event_summary_link,
                          full_report_link,
                          stringsAsFactors = FALSE
                          ))
        }
```

```{r function-getCountries, cache=TRUE, autodep=TRUE}
getCountries <- function(printOnly = FALSE, setGlobalOnly = FALSE) {
        if(!exists("web_not_changed"))
                web_not_changed <- checkIfwebNotChanged()
        if(!web_not_changed)
                warning("The OIE WAHID website has changed, the following may not work.",
                        immediate. = TRUE)
        url <- "http://www.oie.int/wahis_2/public/wahid.php/Countryinformation/Countryhome"
        resp <- GET(url)
        stop_for_status(resp)
        doc <- html(resp, encoding = "UTF-8")
        country_code <- doc %>%
                html_nodes("select#country6 option") %>%
                html_attr("value")
        country_name <- doc %>%
                html_nodes("select#country6 option") %>%
                html_attr("label")
        countries <- data.frame(country_name,
                                country_code,
                                stringsAsFactors = FALSE) %>%
                filter(country_code != "0") %>%
                arrange(country_name)
        class(countries) <- c(class(countries), "OIE searchform countries")
        countries_available <<- countries
        if(setGlobalOnly)
                return(NULL)
        if(printOnly) {
                message("Country codes and names available on the OIE WAHID website:\n")
                message(paste0(names(countries), collapse = ", "))
                message(paste(apply(countries, 1, paste, collapse = ", "),
                                              collapse = "\n"))
                return(NULL)
        }
        return(countries)
}
```

```{r function-getDiseaseSummaries, cache=TRUE, autodep=TRUE}
getDiseaseSummaries <- function(formValues,
                                years,
                                countries,
                                diseaseID,
                                diseaseType) {
        if(!exists("web_not_changed"))
                web_not_changed <- checkIfwebNotChanged()
        if(!web_not_changed)
                warning("The OIE WAHID website has changed, the following may not work.",
                        immediate. = TRUE)
        if(missing(formValues) & any(missing(diseaseID), missing(diseaseType)))
                stop("DiseaseID and type not specified sufficiently.")
        if(!missing(formValues)) {
                if(!"OIE diseaseform values_labels" %in% class(formValues))
                        stop("The argument \"formValues\" is not of class \"OIE diseaseform values_labels\".")
                if(nrow(formValues) > 1)
                        stop(paste0("You specified more than one disease:\n",
                                    paste(formValues[["label"]], collapse = ", ")))
                diseaseID <- formValues[["disease_id_hidden"]]
                diseaseType <- formValues[["disease_type_hidden"]]
                }
        if(!exists("values_labels") | !exists("years_available"))
                getDiseaseFormValues(setGlobalOnly = TRUE)
        diseaseID_Type <- paste(diseaseID, diseaseType, sep = "-")
        diseaseIDs_Types <- apply(values_labels[, c("disease_id_hidden",
                                                    "disease_type_hidden")],
                                  1,
                                  paste0,
                                  collapse = "-")
        if(!diseaseID_Type %in% diseaseIDs_Types)
                stop(paste0("The entered combination of diseaseID and diseaseType (\"",
                            diseaseID_Type,
                            "\") doesn't exist."))
        #         disease <<- formValues[formValues$disease_id_hidden == diseaseID, "label"]
        if(!missing(years)) {
                years <- as.numeric(years)
                if(!all(as.character(years) %in% years_available)) {
                        if(any(as.character(years) %in% years_available)) {
                                warning(paste0("Some of the years entered (",
                                               paste0(years[!as.character(years) %in%
                                                                    years_available],
                                                      collapse = ", "),
                                               ") are outside the range of WAHIS reports, ",
                                               "using only years in the range."),
                                        immediate. = TRUE)
                                years <- years[as.character(years) %in% years_available]
                                } else {
                                        stop(paste0("Year(s) entered (",
                                                    paste0(years, collapse = ", "),
                                                    ") are outside the range of WAHIS reports."))
                                        }
                        }
                } else {
                        message(paste0("Year(s) not specified, using the whole range available (",
                                       paste0(years_available, collapse = ", "),
                                       ")."))
                        years <- years_available
                        }
        if(!missing(countries)) {
                if(!exists("countries_available"))
                        getCountries(setGlobalOnly = TRUE)
                if(!all(countries %in% countries_available$country_name)) {
                        if(any(countries) %in% countries_available$country_name) {
                                warning(paste0("Some of the countries entered (",
                                               paste0(countries[!countries %in%
                                                                        countries_available$country_name],
                                                      collapse = ", "),
                                               ") were not matched in the OIE list of countries, ",
                                               "using only countries that matched."),
                                        immediate. = TRUE)
                                countries <- countries[countries %in% countries_available$country_name]
                                } else {
                                        stop(paste0("Country(ies) entered (",
                                                    paste0(countries, collapse = ", "),
                                                    ") were not matched in the OIE list of countries."))
                                        }
                        } else {
                                countries <- countries
                                }
                
                }
        url <- "http://www.oie.int/wahis_2/public/wahid.php/Diseaseinformation/Immsummary"
        message(paste0("Downloading yearly summaries (", length(years), "):"))
        summaries <- sapply(years,
                            function(x) {
                                    downloadSummaryOfImmediateNotificationsAndFollowUps(year = x,
                                                                                        diseaseID = diseaseID,
                                                                                        diseaseType = diseaseType)
                                    },
                            simplify = FALSE)
        message(paste0("Parsing yearly summaries (", length(summaries), "):"))
        summaries %<>% lapply(parseSummaryOfImmediateNotificationsAndFollowUps)
        for(i in 1:length(summaries)) {
                summaries[[i]] <- cbind(year = years[i], summaries[[i]])
                }
        summaries <- do.call("rbind", summaries)
        if(!missing(countries))
                summaries %<>% filter(country %in% countries)
        summaries %<>% distinct_("reportid")
        class(summaries) %<>% c(., "summary of immediate notifications and followups")
        return(summaries)
        }
```
By inspecting the page source and the forms we figure out the necessary 
parameters to the `getDiseaseSummaries` function: `disease_id_terrestrial` 
(`= "12"` for ASF), and `year`. Now we obtain the yearly summaries of immediate 
notifications and follow-ups as a `summaries` object -- a list containing the raw 
html of pages returned by repeatedly submitting the form with given disease selected 
for the range of years speficied.

```{r getDiseaseSummaries, cache=TRUE, autodep=TRUE}
summaries <- getDiseaseSummaries(formValues = form_values,
                                 years = 2014,
                                 countries = c("Ukraine", "Poland", "Lithuania"))

write.csv2(summaries, file = "summaries.csv", row.names = FALSE, na = "", fileEncoding = "UTF-8")
```

Now we will parse the html of the pages stored in the `summaries` object.

[//]: # (Finish here)

```{r function-downloadSummaryofOutbreaks, cache=TRUE, autodep=TRUE}
downloadSummaryOfOutbreaks <- function(year, reportID, summaryCountry) {
        if(any(missing(year),
               missing(reportID),
               missing(summaryCountry)))
                stop("Not all required arguments specified.")
        url <- "http://www.oie.int/wahis_2/public/wahid.php/Diseaseinformation/Immsummary/listoutbreak"
        if(!exists("web_not_changed"))
                web_not_changed <- checkIfwebNotChanged()
        if(!web_not_changed)
                warning("The OIE WAHID website has changed, the following may not work.",
                        immediate. = TRUE)
        resp <- POST(url = url,
                     body = paste0("reportid=",
                                   reportID,
                                   "&summary_country=",
                                   summaryCountry
                                   ),
                     content_type("application/x-www-form-urlencoded"),
                     add_headers(Referer = "http://www.oie.int/wahis_2/public/wahid.php/Diseaseinformation/Immsummary?reportid=")
                     )
        stop_for_status(resp)
        if(!exists("D2counter")) D2counter <<- 0
        D2counter <<- D2counter + 1
        if(D2counter %% 50 == 0) {
                cat(D2counter, "\n")
                } else {
                        if(D2counter %% 10 == 0) cat("|") else cat(".")
                }
        return(list(year = year,
                    reportid = reportID,
                    summary_country = summaryCountry,
                    resp = resp
                    )
               )
        }
```

```{r function-parseSummaryOfOutbreaks, cache=TRUE, autodep=TRUE}
parseSummaryOfOubreaks <- function(nonParsedSummaryOfOutbreaks) {
        if(!all(c("year",
                  "resp") %in% names(nonParsedSummaryOfOutbreaks)) |
                   !"response" %in% class(nonParsedSummaryOfOutbreaks[["resp"]]))
                stop ("The argument specified is not a list containing a response object.")
        url <- "http://www.oie.int/wahis_2/public/wahid.php/Diseaseinformation/Immsummary/listoutbreak"
        postfields <- rawToChar(nonParsedSummaryOfOutbreaks[["resp"]]
                                [["request"]][["body"]][["config"]][["postfields"]])
        if(!all(nonParsedSummaryOfOutbreaks[["resp"]][["url"]] == url,       
                grep("reportid=", postfields),
                grep("summary_country=", postfields)))
                stop (paste0("The argument specified is not a non parsed summary of outbreaks ", url))
        year <- nonParsedSummaryOfOutbreaks[["year"]]
        doc <- html(nonParsedSummaryOfOutbreaks[["resp"]], encoding = "UTF-8")
        outbreak_country <- doc %>%
                html_nodes(".vacborder:nth-child(1) a") %>%
                html_attr("href") %>%
                gsub("javascript: outbreak_report\\(\\\"|\\\",[0-9]*\\);", "", .)
        outbreak_report <- doc %>%
                html_nodes(".vacborder:nth-child(1) a") %>%
                html_attr("href") %>%
                gsub("javascript: outbreak_report\\(\\\"[A-Z]*\\\",|\\);", "", .)
        full_report_link <- doc %>%
                html_nodes(".vacborder:nth-child(2) a") %>%
                html_attr("href") %>%
                gsub("javascript: open_report\\(\"", "http://www.oie.int", .) %>%
                gsub("\",", "reportid=", .) %>%
                gsub("\\);", "", .)
        loc1 <- doc %>%
                html_nodes(".vacborder:nth-child(3)") %>%
                html_text
        loc2 <- doc %>%
                html_nodes(".vacborder:nth-child(4)") %>%
                html_text
        loc3 <- doc %>%
                html_nodes(".vacborder:nth-child(5)") %>%
                html_text
        epi_unit <- doc %>%
                html_nodes(".vacborder:nth-child(6)") %>%
                html_text
        start_date <- doc %>%
                html_nodes(".vacborder:nth-child(7)") %>%
                html_text %>%
                as.Date("%d/%m/%Y")
        status <- doc %>%
                html_nodes(".vacborder.last") %>%
                html_text
        if(!exists("P2counter")) P2counter <<- 0
        P2counter <<- P2counter + 1
        if(P2counter %% 50 == 0) {
                cat(P2counter, "\n")
                } else {
                        if(P2counter %% 10 == 0) cat("|") else cat(".")
                }
        return(data.frame(year,
                          outbreak_country,
                          outbreak_report,
                          full_report_link,
                          loc1,
                          loc2,
                          loc3,
                          epi_unit,
                          start_date,
                          status,
                          stringsAsFactors = FALSE
                          )
               )
        }

```

```{r function-getSummaryOfOutbreaks, cache=TRUE, autodep=TRUE}
getSummariesOfOutbreaks <- function(diseaseSummaries) {
        if(!"summary of immediate notifications and followups" %in%
                   class(diseaseSummaries))
                stop("The argument \"diseaseSummaries\" is not of class \"summary of immediate notifications and followups\".")
        if(!exists("web_not_changed"))
                web_not_changed <- checkIfwebNotChanged()
        if(!web_not_changed)
                warning("The OIE WAHID website has changed, the following may not work.",
                        immediate. = TRUE)
        message(paste0("Downloading summaries of outbreaks (",
                       nrow(diseaseSummaries),
                       "):"))
        summaries <- apply(diseaseSummaries,
                           1,
                           function(x) downloadSummaryOfOutbreaks(x[["year"]],
                                                               x[["reportid"]],
                                                               x[["summary_country"]]))
        message(paste0("Parsing summaries of outbreaks (", length(summaries), "):"))
        summaries %<>% lapply(parseSummaryOfOubreaks) %>%
                do.call("rbind", .)
        class(summaries) %<>% c(., "summary of outbreaks")
        return(summaries)
}

```

```{r getSummariesOfOutbreaks, cache=TRUE, autodep=TRUE}
summaries_of_outbreaks <- getSummariesOfOutbreaks(summaries)
write.csv2(summaries_of_outbreaks, file = "summaries_of_outbreaks.csv",
            row.names = FALSE,
            na = "",
            fileEncoding = "UTF-8")
```

```{r function-downloadOutbreakReport, cache=TRUE, autodep=TRUE}
downloadOutbreakReport <- function(year,
                                   outbreakReportID,
                                   outbreakCountry,
                                   fullReportID) {
        if(any(missing(year),
               missing(outbreakReportID),
               missing(outbreakCountry),
               missing(fullReportID)))
                stop("Not all required arguments specified.")
        url <- "http://www.oie.int/wahis_2/public/wahid.php/Diseaseinformation/Immsummary/outbreakreport"
        if(!exists("web_not_changed"))
                web_not_changed <- checkIfwebNotChanged()
        if(!web_not_changed)
                warning("The OIE WAHID website has changed, the following may not work.",
                        immediate. = TRUE)
        resp <- POST(url = url,
                     body = paste0("reportid=",
                                   outbreakReportID,
                                   "&summary_country=",
                                   outbreakCountry,
                                   "&backreportid=",
                                   fullReportID
                                   ),
                     content_type("application/x-www-form-urlencoded"),
                     add_headers(Referer = "http://www.oie.int/wahis_2/public/wahid.php/Diseaseinformation/Immsummary/listoutbreak")
                     )
        stop_for_status(resp)
        if(!exists("D3counter")) D3counter <<- 0
        D3counter <<- D3counter + 1
        if(D3counter %% 50 == 0) {
                cat(D3counter, "\n")
                } else {
                        if(D3counter %% 10 == 0) cat("|") else cat(".")
                }
        return(list(year = year,
                    reportid = outbreakReportID,
                    summary_country = outbreakCountry,
                    resp = resp))
        }
```

```{r function-parseOutbreakReport, cache=TRUE, autodep=TRUE}
parseOutbreakReport <- function(nonParsedOutbreakReport) {
        if(!all(c("year",
                  "summary_country",
                  "reportid",
                  "resp") %in% names(nonParsedOutbreakReport)) |
                   !"response" %in% class(nonParsedOutbreakReport[["resp"]]))
                stop ("The argument specified is not a list containing a response object.")
        url <- "http://www.oie.int/wahis_2/public/wahid.php/Diseaseinformation/Immsummary/outbreakreport"
        postfields <- rawToChar(nonParsedOutbreakReport[["resp"]]
                                [["request"]][["body"]][["config"]][["postfields"]])
        if(!all(nonParsedOutbreakReport[["resp"]][["url"]] == url,       
                grep("reportid=", postfields),
                grep("summary_country=", postfields),
                grep("backreportid=", postfields)))
                stop (paste0("The argument specified is not a non parsed outbreak report ", url))
        year <- nonParsedOutbreakReport[["year"]]
        outbreak_country <- nonParsedOutbreakReport[["summary_country"]]
        reportid <- nonParsedOutbreakReport[["reportid"]]
        doc <- html(nonParsedOutbreakReport[["resp"]], encoding = "UTF-8")
        data <- doc %>%
                html_nodes(".vacborder") %>%
                html_text
        start_date <- data[1] %>% as.Date("%d/%m/%Y")
        status <- data[2]
        resolution_date <- data[3] %>% as.Date("%d/%m/%Y")
        loc1 <- data[4]
        loc2 <- data[5]
        loc3 <- data[6]
        loc4 <- data[8]
        epi_unit <- data[7]
        lat <- data[9] %>% as.numeric
        lon <- data[10] %>% as.numeric
        description <- data[11]
        at_risk_total <- data[length(data) - 3] %>% as.numeric
        cases_total <- data[length(data) - 2] %>% as.numeric
        deaths_total <- data[length(data) - 1] %>% as.numeric
        destroyed_total <- data[length(data)] %>% as.numeric
        multi_species <- length(data) > 20
        species <- paste0(data[seq(12, length(data) - 4, 5)], collapse = ", ") # ERROR HERE
        species_breakdown <- data[12:(length(data) - 4)] %>%
                matrix(ncol = 5,
                       byrow = TRUE,
                       dimnames = list(NULL,
                                       c("species",
                                         "at_risk",
                                         "cases",
                                         "deaths",
                                         "destroyed"))) %>%
                data.frame(stringsAsFactors = FALSE)
        species_breakdown[, 2:5] <- as.numeric(
                as.character(unlist(species_breakdown[, 2:5])))
        ret <- data.frame(outbreak_country,
                          reportid,
                          start_date,
                          status,
                          resolution_date,
                          loc1,
                          loc2,
                          loc3,
                          loc4,
                          epi_unit,
                          lat,
                          lon,
                          description,
                          at_risk_total,
                          cases_total,
                          deaths_total,
                          destroyed_total,
                          multi_species,
                          species,
                          species_breakdown,
                          stringsAsFactors = FALSE,
                          row.names = NULL)
        if(!exists("P3counter")) P3counter <<- 0
        P3counter <<- P3counter + 1
        if(P3counter %% 50 == 0) {
                cat(P3counter, "\n")
                } else {
                        if(P3counter %% 10 == 0) cat("|") else cat(".")
                }
        return(ret)        
        }
```

```{r function-getOutbreakReports, cache=TRUE, autodep=TRUE}
getOutbreakReports <- function(outbreakSummaries, writeCSV = FALSE) {
        if(!"summary of outbreaks" %in% class(outbreakSummaries))
                stop("The argument \"outbreakSummaries\" is not of class \"summary of outbreaks\".")
        if(!exists("web_not_changed"))
                web_not_changed <- checkIfwebNotChanged()
        if(!web_not_changed)
                warning("The OIE WAHID website has changed, the following may not work.",
                        immediate. = TRUE)
        message(paste0("Downloading outbreak reports (",
                       nrow(outbreakSummaries),
                       "):"))
        reports <- apply(outbreakSummaries,
                         1,
                         function(x) downloadOutbreakReport(x[["year"]],
                                                            x[["outbreak_report"]],
                                                            x[["outbreak_country"]],
                                                            gsub("http://www.oie.int/wahis_2/public/wahid.php/Reviewreport/Review\\?page_refer=MapFullEventReport&reportid=",
                                                                 "",
                                                                 x[["full_report_link"]])))
        message(paste0("Parsing outbreak reports(", length(reports), "):"))
        reports %<>%
                lapply(parseOutbreakReport) %>%
                do.call("rbind", .)
        if(writeCSV) {
                write.table(reports,
                          file = paste0("outbreaks of ",
                                        disease,
                                        ", year ",
                                        paste0(unique(range(years)), collapse = "-"),
                                        ", countries",
                                        paste0(countries, collapse = ", "),
                                        ".txt"),
                          na = "",
                          fileEncoding = "UTF-8")
                return(NULL)
                } else {
                        return(reports)
                        }
        }
```

```{r getOutbreakReports, cache=TRUE, autodep=TRUE}
outbreak_reports <- getOutbreakReports(summaries_of_outbreaks)
write.csv2(outbreak_reports, file = "outbreak_reports.csv", row.names = FALSE,
           na = "", fileEncoding = "UTF-8")
```

```{r examplePlot, cache=TRUE, autodep=TRUE}
# library(plotGoogleMaps)
# outbreak_reportsSPDF <- outbreak_reports
# coordinates(outbreak_reportsSPDF) <- ~lon + lat
# proj4string(outbreak_reportsSPDF) <- CRS("+proj=longlat +datum=WGS84")
# ic <- iconlabels(outbreak_reportsSPDF$species, height = 5, icon = TRUE)
# map <- plotGoogleMaps(outbreak_reportsSPDF,
#                       file = "map.html",
#                       zcol = "start_date",
#                       mapTypeId = "TERRAIN",
#                       legend = FALSE,
#                       control = FALSE,
#                       iconMarker = ic)

# library(ggmap)
# theme_set(theme_bw(16))
# europe_map <- qmap("Ukraine", zoom = 5, source = "stamen")
# europe_map +
#         geom_point(aes(x = lon, y = lat, colour = start_date, size = cases_total),
#                    data = outbreak_reports)
# qmplot(lon, lat, data = outbreak_reports)

```

Cite packages!
Encodings!
Specify a country!
Make a consistent flow!
