---
title: Making a map of African Swine Fever (ASF) outbreaks in Europe from the data
  scrapped from the OIE WAHID website
author: "Pavel Vodrážka"
date: "Thursday, January 15, 2015"
output:
  html_document: default
---

### Reason

The maps that can be obtained from the [OIE WAHID website](http://www.oie.int/wahis_2/public/wahid.php/Diseaseinformation/Diseaseoutbreakmaps) are not aesthetically 
pleasing, and can be downloaded only as raster images that are not suitable for 
including in print quality publications. Other sources of data and maps of ASF 
outbreaks were not found.

### Obtaining the data

There is no option to dowload the data or no public data API on the OIE WAHID 
website. Data needs to be scraped from the ["Immediate notifications and 
Follow-Ups"](http://www.oie.int/wahis_2/public/wahid.php/Diseaseinformation/Immsummary) part of the site.

We go all the way from getting the webpages of summaries of immediate 
notifications and follow-ups for african swine fever for all the years (2005 -- 
2014), extracting the links of the full country reports, getting the full 
reports (immediate notifications), extracting the links of the follow-up 
reports, getting the follow-up reports, and getting and processing data from the 
dowloaded reports.

The starting page is http://www.oie.int/wahis_2/public/wahid.php/Diseaseinformation/Immsummary. The following code refers to the version of the web denoted in the 
footer:
```
+--------------------------------------------------------------+
| World Animal Health Information Database (WAHID) – Version 1 |
| Copyright © World Organisation for Animal Health (OIE)       |
| Release date: 30 December 2013 August 2012                   |
+--------------------------------------------------------------+
```
```{r loadPackages}
library(rvest)
library(httr)
library(dplyr)
library(magrittr)
library(digest)
library(rworldmap)
```

```{r function-checkIfwebNotChanged, cache=TRUE}
checkIfwebNotChanged <- function(hashFile = "hashes.rds") {
        # do file opening with error catching
        message("Checking whether the OIE WAHID website structure didn't change")
        url <- "http://www.oie.int/wahis_2/public/wahid.php/Diseaseinformation/Immsummary"
        val <- list()
        resp1 <- GET(url)
        stop_for_status(resp1)
        doc1 <- html(resp1, encoding = "UTF-8")
        val[["disease form attributes"]] <- doc1 %>%
                html_nodes("#diseaseform") %>%
                html_attrs %>%
                digest(algo = "md5")
        val[["diseaseform scripts"]] <- doc1 %>%
                html_nodes("#diseaseform script") %>%
                digest(algo = "md5")
        val[["diseaseform inputs"]] <- doc1 %>%
                html_nodes("#diseaseform input") %>%
                digest(algo = "md5")
        val[["diseaseform selects' attributes"]] <- doc1 %>%
                html_nodes("select") %>%
                html_attrs %>%
                digest(algo = "md5")
        resp2 <- GET(paste0(url, "/listoutbreak"))
        stop_for_status(resp2)
        doc2 <- html(resp2, encoding = "UTF-8")
        val[["outbreakreport form attributes"]] <- doc2 %>%
                html_nodes("form[name='outbreakreport']") %>%
                html_attrs %>%
                digest(algo = "md5")
        val[["outbreakreport form inputs"]] <- doc2 %>%
                html_nodes("form[name='outbreakreport'] input") %>%
                digest(algo = "md5")
        
        if(file.exists(hashFile)) {
                saved <- readRDS(hashFile)
                comparison <- mapply(identical, val, saved)
                if(!all(comparison)) {
                        not_equal <- names(comparison[!comparison])
                        warning("There are changes on the OIE WAHID website ", 
                                "that can affect the functionality of this package.",
                                "\nThings that have changed are:\n- ",
                                paste0(not_equal, collapse = "\n- "),
                                immediate. = TRUE)
                        web_not_changed <<- FALSE
                        return(FALSE)
                        } else {
                                web_not_changed <<- TRUE
                                message("... OK")
                                return(TRUE)
                                }
                } else {
                        warning("Impossible to check whether the OIE WAHID website ", 
                                "has not changed (hash file not present, creating ",
                                "a new one from the current state of the site).",
                                immediate. = TRUE)
                        saveRDS(val, hashFile)
                        web_not_changed <<- TRUE
                        return(TRUE)
                        }
        }       
```

We first need to be able to programmatically fill the form on the page and download 
the resulting summaries. We prepare a function `getDiseaseSummaries` using some 
functions from the "RHTMLForms" package to accomplish this part of the task.

There are four forms on the page; after inspecting the page source we know that 
we are interested in the `diseaseform`. Our function `getDiseaseSummaries` 
contains a function `summaries_fun` that is created programmatically from the 
page source code and that will serve to submit that form and retrieve the 
resulting webpages.

```{r function-getDiseaseFormValues, cache=TRUE, autodep=TRUE}
getDiseaseFormValues <- function(disease = character(),
                                 printOnly = FALSE,
                                 setGlobalOnly = FALSE,
                                 newDownload = FALSE,
                                 file = "diseaseFormValues.rds") {
        message("Getting the form values for disease(s) entered")
        if(!exists("web_not_changed"))
                web_not_changed <- checkIfwebNotChanged()
#         if(!web_not_changed)
#                 warning("The OIE WAHID website has changed, the following may not work.",
#                         immediate. = TRUE)
        if(file.exists(file) & !newDownload) {
                current_year <- as.integer(format(Sys.Date(), "%Y"))
                cached <- readRDS(file)
                values_labels <<- cached[["values_labels"]]
                if(!"OIE diseaseform values_labels" %in% class(values_labels))
                        newDownload <- TRUE
                years_available <<- cached[["years_available"]]
                if(!all(range(as.integer(years_available)) == c(2005, current_year)))
                        newDownload <- TRUE
                }
        if((!file.exists(file)) | newDownload) {
                message("... downloading")
                url <- "http://www.oie.int/wahis_2/public/wahid.php/Diseaseinformation/Immsummary"
                resp <- GET(url)
                stop_for_status(resp)
                doc <- html(resp, encoding = "UTF-8")
                years_available <<- doc %>%
                        html_nodes("#diseaseform #year option") %>%
                        html_attr("value")
                terrestrial_values <- doc %>%
                        html_nodes("#diseaseform #disease_id_terrestrial option") %>%
                        html_attr("value")
                terrestrial_labels <- doc %>%
                        html_nodes("#diseaseform #disease_id_terrestrial option") %>%
                        html_attr("label")
                aquatic_values <- doc %>%
                        html_nodes("#diseaseform #disease_id_aquatic option") %>%
                        html_attr("value")
                aquatic_labels <- doc %>%
                        html_nodes("#diseaseform #disease_id_aquatic option") %>%
                        html_attr("label")
                values_labels <<- data_frame(label = c(terrestrial_labels, aquatic_labels),
                                             disease_id_hidden = c(terrestrial_values,
                                                                   aquatic_values),
                                             disease_type_hidden = c(rep.int("0",
                                                                             length(terrestrial_values)),
                                                                     rep.int("1",
                                                                             length(aquatic_values))
                                                                     )) %>%
                        filter(disease_id_hidden != "-999") %>%
                        arrange(label)
                class(values_labels) <- c(class(values_labels), "OIE diseaseform values_labels")
                saveRDS(list(values_labels = values_labels,
                             years_available = years_available),
                        file)
                } else {
                        message("... using cached data")
                }
        if(setGlobalOnly)
                return(invisible(NULL))
        if(length(disease) > 0) {
                values_labels_filtered <- values_labels[grep(disease,
                                                             values_labels$label,
                                                             ignore.case = TRUE)
                                                        ,]
                class(values_labels_filtered) <- c(class(values_labels_filtered),
                                                   "OIE diseaseform values_labels")
                }
        if(printOnly) {
                message("Values for submitting the \"diseaseform\" on the OIE WAHID ", 
                        "\"Summary of Immediate notifications and Follow-ups\" ",
                        "webpage\n[http://www.oie.int/wahis_2/public/wahid.php/",
                        "Diseaseinformation/Immsummary)]:\n")
                if(length(disease) > 0) {
                        message("Entered search for disease: \"", disease, "\"\n")
                        if(nrow(values_labels_filtered) == 0) {
                                message("No disease found. Diseases available: \n")
                                message(paste0(names(values_labels), collapse = ", "))
                                message(paste(apply(values_labels, 1, paste, collapse = ", "),
                                              collapse = "\n"))
                                } else {
                                        message(paste0(names(values_labels_filtered), collapse = ", "))
                                        message(paste(apply(values_labels_filtered, 1, paste, collapse = ", "),
                                                      collapse = "\n"))
                                        }
                        } else {
                                message(paste0(names(values_labels), collapse = ", "))
                                message(paste(apply(values_labels, 1, paste, collapse = ", "),
                                              collapse = "\n"))
                                }
                message("\nYears available:\n")
                message(paste0(range(years_available), collapse = " - "))
                return(invisible(NULL))
                } else {
                        if(length(disease > 0)) {
                                if(nrow(values_labels_filtered) > 1)
                                        warning("Your search for pattern \"",
                                                disease,
                                                "\" returned more than one disease.",
                                                immediate. = TRUE)
                                if(nrow(values_labels_filtered) == 0) {
                                        warning("Your search for pattern \"",
                                                disease,
                                                "\" returned no disease.",
                                                immediate. = TRUE)
                                        return(NULL)
                                        }
                                return(values_labels_filtered)
                                } else {
                                        return(values_labels)
                                        }        
                        }
        }
```

```{r obtainFormValues, cache=TRUE, autodep=TRUE}
form_values <- getDiseaseFormValues(disease = "african swine fever")
```

```{r function-downloadSINF, cache=TRUE, autodep=TRUE}
downloadSINF <- function(disease_id_hidden,
                         disease_type_hidden,
                         year
                         ) {
        if(any(missing(disease_id_hidden),
               missing(disease_type_hidden),
               missing(year)))
                stop("Not all required arguments specified.")
        if(!exists("web_not_changed"))
                web_not_changed <- checkIfwebNotChanged()
#         if(!web_not_changed)
#                 warning("The OIE WAHID website has changed, the following may not work.",
#                         immediate. = TRUE)
        if(!exists("values_labels") | !exists("years_available"))
                getDiseaseFormValues(setGlobalOnly = TRUE)
        diseaseID_Type <- paste(disease_id_hidden, disease_type_hidden, sep = "-")
        diseaseIDs_Types <- apply(values_labels[, c("disease_id_hidden",
                                                    "disease_type_hidden")],
                                  1,
                                  paste0,
                                  collapse = "-")
        if(!diseaseID_Type %in% diseaseIDs_Types)
                stop("The entered combination of disease_id_hidden ",
                     "and disease_type_hidden (\"",
                     diseaseID_Type,
                     "\") doesn't exist.")
        if(!as.character(year) %in% years_available)
                stop("The year entered (\"",
                     year,
                     "\") is outside the range of WAHIS reports.")
        url <- "http://www.oie.int/wahis_2/public/wahid.php/Diseaseinformation/Immsummary"
        resp <- POST(url = url,
                     body = list(disease_id_hidden = disease_id_hidden,
                                 disease_type_hidden = disease_type_hidden,
                                 year = year),
                     encode = "form")
        stop_for_status(resp)
        if(!exists("D1counter")) D1counter <<- 0
        D1counter <<- D1counter + 1
        if(D1counter %% 50 == 0) {
                cat(D1counter, "\n")
                } else {
                        if(D1counter %% 10 == 0) cat("|") else cat(".")
                }
        return(list(year = year,
                    disease_id_hidden = disease_id_hidden,
                    disease_type_hidden = disease_type_hidden,
                    resp = resp))
        }
```

```{r function-parseSINF, cache=TRUE, autodep=TRUE}
parseSINF <- function(nonParsedSINF) {
        if(!all(c("year",
                  "disease_id_hidden",
                  "disease_type_hidden",
                  "resp") %in% names(nonParsedSINF)) |
                   !"response" %in% class(nonParsedSINF[["resp"]]))
                stop ("The argument specified is not a list containing a response object.")
        response <- nonParsedSINF[["resp"]]
        url <- "http://www.oie.int/wahis_2/public/wahid.php/Diseaseinformation/Immsummary"
        postfields <- rawToChar(response[["request"]][["body"]][["config"]][["postfields"]])
        if(!all(response["url"] == url,       
                grep("disease_id_hidden=", postfields),
                grep("disease_type_hidden=", postfields),
                grep("year=", postfields)
                )) stop ("The argument specified is not a summary ",
                         "of immediate notifications and follow-ups ",
                         url)
        year <- nonParsedSINF[["year"]]
        disease_id_hidden <- nonParsedSINF[["disease_id_hidden"]]
        disease_type_hidden <- nonParsedSINF[["disease_type_hidden"]]
        doc <- html(response, encoding = "UTF-8")
        country <- doc %>%
                html_nodes(".outbreakdetails .outbreak_country") %>%
                html_text(trim = TRUE) %>%
                gsub("[^A-Za-z ']", "", .)
        while(length((empty_countries <- which(nchar(country) == 0))) > 0) {
                country[empty_countries] <- country[empty_countries - 1]
                }
        status <- doc %>%
                html_nodes(".outbreakdetails :nth-child(2)") %>%
                html_text(trim = TRUE) %>%
                gsub("[0-9/]", "", .)
        date <- doc %>%
                html_nodes(".outbreakdetails :nth-child(2)") %>%
                html_text(trim = TRUE) %>%
                gsub("[A-Za-z ]", "", .) %>%
                as.Date("%d/%m/%Y")
        summary_country <- doc %>%
                html_nodes("#diseaseform :nth-child(3) a") %>%
                html_attr("onclick") %>%
                gsub("outbreaklist\\('|',[0-9]*);", "", .)
        reportid <- doc %>%
                html_nodes("#diseaseform :nth-child(3) a") %>%
                html_attr("onclick") %>%
                gsub("outbreaklist\\('[A-Z]*',|);", "", .)
        event_summary_link <- doc %>%
                html_nodes(".vacborder:nth-child(4) a") %>%
                html_attr("href") %>%
                paste0("http://www.oie.int", .)
        full_report_link <- doc %>%
                html_nodes(".vacborder:nth-child(5) a") %>%
                html_attr("href") %>%
                gsub("javascript: open_report\\(\"", "http://www.oie.int", .) %>%
                gsub("\",", "reportid=", .) %>%
                gsub("\\);", "", .)
        if(!exists("P1counter")) P1counter <<- 0
        P1counter <<- P1counter + 1
        if(P1counter %% 50 == 0) {
                cat(P1counter, "\n")
                } else {
                        if(P1counter %% 10 == 0) cat("|") else cat(".")
                }
        return(data_frame(year,
                          disease_id_hidden,
                          disease_type_hidden,
                          country = if(length(full_report_link) > 0) country else NA,
                          status = if(length(full_report_link) > 0) status else NA,
                          date = if(length(full_report_link) > 0) date else NA,
                          summary_country = if(length(full_report_link) > 0) summary_country else NA,
                          reportid = if(length(full_report_link) > 0) reportid else NA,
                          event_summary_link = if(length(full_report_link) > 0) event_summary_link else NA,
                          full_report_link = if(length(full_report_link) > 0) full_report_link else NA))
        }
```

```{r function-getCountries, cache=TRUE, autodep=TRUE}
getCountries <- function(printOnly = FALSE,
                         setGlobalOnly = FALSE,
                         newDownload = FALSE,
                         file = "countriesAvailable.rds") {
        message("Getting available countries")
        if(!exists("web_not_changed"))
                web_not_changed <- checkIfwebNotChanged()
#         if(!web_not_changed)
#                 warning("The OIE WAHID website has changed, the following may not work.",
#                         immediate. = TRUE)
        if(file.exists(file) & !newDownload) {
                countries_available <<- readRDS(file)
                if(!"OIE searchform countries" %in% class(countries_available))
                        newDownload <- TRUE
                }
        if(!file.exists(file) | newDownload) {
                message("... downloading")
                url <- "http://www.oie.int/wahis_2/public/wahid.php/Countryinformation/Countryhome"
                resp <- GET(url)
                stop_for_status(resp)
                doc <- html(resp, encoding = "UTF-8")
                continent_code <- doc %>%
                        html_nodes("select#locn1 option") %>%
                        html_attr("value")
                continent_name <- doc %>%
                        html_nodes("select#locn1 option") %>%
                        html_attr("label")
                continents <- data_frame(continent_name,
                                         continent_code) %>%
                        filter(continent_code != "0",
                               continent_name != "Entire World")
                data(countryRegions, package = "rworldmap")
                countries <- apply(continents, 1, function(x) {
                        ISO3 <- doc %>%
                                html_nodes(paste0("select#country",
                                                  x[2],
                                                  " option")) %>%
                                html_attr("value")
                        OIE_name <- doc %>%
                                html_nodes(paste0("select#country",
                                                  x[2],
                                                  " option")) %>%
                                html_attr("label")
                        return(data_frame(ISO3,
                                          OIE_name,
                                          OIE_region = x[1]))
                        
                        }) %>%
                        do.call(rbind, .) %>%
                        left_join(select(countryRegions,
                                         ISO3:GEO3))
                class(countries) <- c(class(countries), "OIE searchform countries")
                countries_available <<- countries
                saveRDS(countries_available, file)
                } else {
                        message("... using cached data")
                }
        if(setGlobalOnly)
                return(invisible(NULL))
        if(printOnly) {
                message("Country codes and names available on the OIE WAHID website:\n")
                message(paste0(names(countries), collapse = ", "))
                message(paste(apply(countries, 1, paste, collapse = ", "),
                                              collapse = "\n"))
                return(invisible(NULL))
        }
        return(countries)
}
```

```{r function-getSINF, cache=TRUE, autodep=TRUE}
getSINF <- function(formValues,
                    years,
                    ISO3Codes,
                    countryNames,
                    OIERegions,
                    regions,
                    continents,
                    GEO3Majors,
                    GEO3s,
                    diseaseID,
                    newDownload = FALSE,
                    file = "SINF.rds") {
        message("Getting summaries of immediate notifications and follow-ups")
        if(missing(formValues) & missing(diseaseID))
                stop("Disease not specified.")
        loc_arg_spec <- sum(c(!missing(ISO3Codes),
                              !missing(countryNames),
                              !missing(OIERegions),
                              !missing(regions),
                              !missing(continents),
                              !missing(GEO3Majors),
                              !missing(GEO3s)))
        if(loc_arg_spec > 1)
                stop("Multiple specification of countries.")
        if(!exists("web_not_changed"))
                web_not_changed <- checkIfwebNotChanged()
#         if(!web_not_changed)
#                 warning("The OIE WAHID website has changed, the following may not work.",
#                         immediate. = TRUE)
        if(!exists("values_labels") | !exists("years_available"))
                getDiseaseFormValues(setGlobalOnly = TRUE)
        if(!missing(formValues)) {
                if(!"OIE diseaseform values_labels" %in% class(formValues))
                        stop("The argument \"formValues\" is not of class ",
                             "\"OIE diseaseform values_labels\".")
                if(nrow(formValues) > 1)
                        stop("You specified more than one disease:\n",
                             paste(formValues[["label"]], collapse = ", "))
                diseaseID <- formValues[["disease_id_hidden"]]
                diseaseType <- formValues[["disease_type_hidden"]]
                diseaseID_Type <- paste(diseaseID, diseaseType, sep = "-")
                diseaseIDs_Types <- apply(select(values_labels, -label),
                                          1,
                                          paste0,
                                          collapse = "-")
                if(!diseaseID_Type %in% diseaseIDs_Types)
                        stop("The entered combination of diseaseID and diseaseType (\"",
                             diseaseID_Type,
                             "\") doesn't exist.")
                } else {
                        if(missing(diseaseID)) {
                            stop("Disease not specified.")    
                        }
                        diseaseType <- filter(values_labels,
                                              disease_id_hidden = diseaseID)
                        if(length(diseaseType) == 0)
                                stop("The entered diseaseID doesn't exist.")
                }
        disease <- select(filter(values_labels,
                                 disease_id_hidden == diseaseID),
                          label)
        if(!missing(years)) {
                years <- as.character(years)
                if(!all(years %in% years_available)) {
                        if(any(years %in% years_available)) {
                                warning("Some of the years entered (",
                                        paste0(years[!years %in%
                                                             years_available],
                                               collapse = ", "),
                                        ") are outside the range of WAHIS reports, ",
                                        "using only years in the range.",
                                        immediate. = TRUE)
                                years <- years[years %in% years_available]
                                } else {
                                        stop("Year(s) entered (",
                                             paste0(years, collapse = ", "),
                                             ") are outside the range of WAHIS reports.")
                                        }
                        }
                } else {
                        message("- Year(s) not specified, using the whole range available (",
                                paste0(years_available, collapse = ", "),
                                ").")
                        years <- years_available
                        }
        if(!exists("countries_available"))
                getCountries(setGlobalOnly = TRUE)
        if(loc_arg_spec == 1) {
                if(!missing(countryNames)) {
                        not_matched <- countryNames[sapply(countryNames,
                               function(y) all(apply(countries_available,
                                                     1,
                                                     function(x) length(grep(y, x, invert = TRUE)) == 8)))]
                        if(length(not_matched) == length(countryNames)) {
                                stop("Countries or regions entered (countryNames = ",
                                     deparse(countryNames),
                                     ") were not matched in the OIE list of countries.")
                                }
                        if(length(not_matched) > 0) {
                                warning("Some of the countries or regions entered (countryNames = ",
                                        deparse(not_matched),
                                        ") were not matched in the OIE list of countries, ",
                                        "using only countries that matched.",
                                        immediate. = TRUE)
                                }
                        countries <- lapply(countryNames,
                                            function(y) countries_available[apply(
                                                    countries_available,
                                                    1,
                                                    function(x) sum(grep(y, x)) > 0), ])
                        message("- Countries that matched to your search terms:")
                        names(countries) <- countryNames
                        lapply(names(countries), function(x) {
                                message(" \"", x, "\": ",
                                        paste0(countries[[x]][["OIE_name"]],
                                               collapse = ", "))})
                        countries %<>%
                                do.call("rbind", .)
                        } else {
                                if(!missing(ISO3Codes)) {
                                        sel <- ISO3Codes
                                        col <- "ISO3"
                                        }        
                                
                                if(!missing(OIERegions)) {
                                        sel <- OIERegions
                                        col <- "OIE_region"
                                        }
                                if(!missing(regions)) {
                                        sel <- regions
                                        col <- "REGION"
                                        }
                                if(!missing(continents)) {
                                        sel <- continents
                                        col <- "continent"
                                        }
                                if(!missing(GEO3Majors)) {
                                        sel <- GEO3Majors
                                        col <- "GEO3major"
                                        }
                                if(!missing(GEO3s)) {
                                        sel <- GEO3s
                                        col <- "GEO3"
                                        }
                                if(!all(sel %in% unlist(countries_available[, col]))) {
                                        matched <- sel[sel %in% unlist(countries_available[, col])]
                                        if(length(matched) == 0) {
                                                stop("Countries or regions entered (",
                                                     col,
                                                     " = ",
                                                     deparse(sel),
                                                     ") were not matched in the OIE list of countries.")
                                                } else {
                                                        warning("Some of the countries or regions entered (",
                                                                col,
                                                                " = ",
                                                                deparse(sel[!sel %in%
                                                                                    unlist(countries_available[, col])]),
                                                                ") were not matched in the OIE list of countries, ",
                                                                "using only countries that matched.",
                                                                immediate. = TRUE)
                                                        }
                                        }
                                countries <- countries_available[unlist(countries_available[, col]) %in% sel, ]
                                }
                } else {
                        countries <- countries_available
                        }
        templateSINFs <- data_frame(disease = character(0),
                                    year = character(0),
                                    disease_id_hidden = character(0),
                                    disease_type_hidden = character(0),
                                    country = character(0),
                                    status = character(0),
                                    date = as.Date(character(0),
                                                   format = "%Y-%m-%d"),
                                    summary_country = character(0),
                                    reportid = character(0),
                                    event_summary_link = character(0),
                                    full_report_link = character(0))
        entered <- data_frame(disease_id_hidden = diseaseID, year = years)
        message("- Entered (", nrow(entered), "):")
        print(entered)
        if(file.exists(file) & !newDownload) {
                message("- Using cache file.")
                cachedSINFs <- readRDS(file)
                if(!"summary of immediate notifications and followups"
                   %in% class(cachedSINFs)) {
                        warning("Cached SINFs file ignored because it doesn't ",
                                "have right format, building a new one.",
                                immediate. = TRUE)
                        cachedSINFs <- templateSINFs
                        class(cachedSINFs) %<>% c(.,
                                                  "summary of immediate notifications and followups")
                        }
                } else {
                        cachedSINFs <- templateSINFs
                        class(cachedSINFs) %<>% c(., "summary of immediate notifications and followups")
                        }
        suppressMessages(cached <- inner_join(cachedSINFs, entered))
        message("- Cached (", nrow(cached),"):")
        print(cached)
        to_download <- years[!years %in% cached[["year"]]]
        message("- To download (", length(to_download), "):")
        print(to_download)
        if(!newDownload) {
                suppressMessages(back_to_cache <- entered %>%
                                         anti_join(cachedSINFs, .) %>%
                                         rbind(cached))
                } else {
                        suppressMessages(back_to_cache <- anti_join(cachedSINFs,
                                                                    entered))
                        }
        message("- Back to cache (", nrow(back_to_cache), "):")
        print(back_to_cache)
        url <- "http://www.oie.int/wahis_2/public/wahid.php/Diseaseinformation/Immsummary"
        message("- Downloading yearly summaries (", length(to_download), "):")
        D1counter <<- 0
        summaries <- sapply(to_download,
                            function(x) {
                                    downloadSINF(year = x,
                                                 disease_id_hidden = diseaseID,
                                                 disease_type_hidden = diseaseType)
                                    },
                            simplify = FALSE)
        message("- Parsing yearly summaries (", length(summaries), "):")
        P1counter <<- 0
        summaries %<>% lapply(parseSINF)
        summaries <- do.call("rbind", summaries)
        summaries <- as_data_frame(c(list(disease = rep(as.character(disease),
                                                        nrow(summaries))),
                                     as.list(summaries)))
#         attr(summaries, "disease") <- disease
#         attr(summaries, "countries") <- countries[["ISO3"]]
#         attr(summaries, "years") <- years
        summaries_to_write <- summaries[!(summaries$year %in%
                                                  unlist(summaries[summaries$status == "Continuing",
                                                                   "year"]))
                                        & summaries$year != as.integer(format(Sys.Date(), "%Y")), ]
        message("- Downloaded to write (", nrow(summaries_to_write), "):")
        print(summaries_to_write)
        write_cache <- rbind(back_to_cache,
                             summaries_to_write)
        class(write_cache) %<>% c(., "summary of immediate notifications and followups")
        saveRDS(write_cache, file)
        summaries_to_return <- rbind(cached, summaries)
        if(!missing(countries))
                summaries_to_return %<>% filter(summary_country %in% countries[["ISO3"]])
        summaries_to_return %<>% distinct_("reportid")
        class(summaries_to_return) %<>% c(., "summary of immediate notifications and followups")
        return(summaries_to_return)
        }
```

By inspecting the page source and the forms we figure out the necessary 
parameters to the `getDiseaseSummaries` function: `disease_id_terrestrial` 
(`= "12"` for ASF), and `year`. Now we obtain the yearly summaries of immediate 
notifications and follow-ups as a `summaries` object -- a list containing the raw 
html of pages returned by repeatedly submitting the form with given disease selected 
for the range of years speficied.

```{r getDiseaseSummaries, cache=TRUE, autodep=TRUE}
summaries <- getSINF(formValues = form_values
                     #                                  ,years = 2014
                     #                                 ,countryNames = "Africa"
                     #                                 ,newDownload = TRUE
                     )

write.csv2(summaries, file = "summaries.csv", row.names = FALSE, na = "", fileEncoding = "UTF-8")
```

Now we will parse the html of the pages stored in the `summaries` object.

[//]: # (Finish here)

```{r function-downloadSO, cache=TRUE, autodep=TRUE}
downloadSO <- function(summaryRecord) {
        if(!all(c("reportid", "summary_country") %in% names(summaryRecord)))
                stop("Not all required arguments specified.")
        url <- "http://www.oie.int/wahis_2/public/wahid.php/Diseaseinformation/Immsummary/listoutbreak"
        if(!exists("web_not_changed"))
                web_not_changed <- checkIfwebNotChanged()
#         if(!web_not_changed)
#                 warning("The OIE WAHID website has changed, the following may not work.",
#                         immediate. = TRUE)
        resp <- POST(url = url,
                     body = list(reportid = summaryRecord[["reportid"]],
                                 summary_country = summaryRecord[["summary_country"]]),
                     encoding = "form",
                     add_headers(Referer = "http://www.oie.int/wahis_2/public/wahid.php/Diseaseinformation/Immsummary?reportid="))
        stop_for_status(resp)
        if(!exists("D2counter")) D2counter <<- 0
        D2counter <<- D2counter + 1
        if(D2counter %% 50 == 0) {
                cat(D2counter, "\n")
                } else {
                        if(D2counter %% 10 == 0) cat("|") else cat(".")
                }
        return(c(summaryRecord, list(resp = resp)))
        }
```

```{r function-parseSO, cache=TRUE, autodep=TRUE}
parseSO <- function(x) {
        if(!all(c("resp") %in% names(x)) |
                   !"response" %in% class(x[["resp"]]))
                stop ("The argument specified is not a list containing a response object.")
        url <- "http://www.oie.int/wahis_2/public/wahid.php/Diseaseinformation/Immsummary/listoutbreak"
        if(!all(x[["resp"]][["url"]] == url,       
                grep("reportid", names(x[["resp"]][["request"]][["body"]][["body"]])),
                grep("summary_country", names(x[["resp"]][["request"]][["body"]][["body"]]))))
                stop ("The argument specified is not a non parsed summary of outbreaks ", url)        
        doc <- html(x[["resp"]], encoding = "UTF-8")
        x[["outbreak_country"]] <- doc %>%
                html_nodes(".vacborder:nth-child(1) a") %>%
                html_attr("href") %>%
                gsub("javascript: outbreak_report\\(\\\"|\\\",[0-9]*\\);", "", .)
        x[["outbreak_report"]] <- doc %>%
                html_nodes(".vacborder:nth-child(1) a") %>%
                html_attr("href") %>%
                gsub("javascript: outbreak_report\\(\\\"[A-Z]*\\\",|\\);", "", .)
        x[["full_report_link"]] <- doc %>%
                html_nodes(".vacborder:nth-child(2) a") %>%
                html_attr("href") %>%
                gsub("javascript: open_report\\(\"", "http://www.oie.int", .) %>%
                gsub("\",", "reportid=", .) %>%
                gsub("\\);", "", .)
        x[["loc1"]] <- doc %>%
                html_nodes(".vacborder:nth-child(3)") %>%
                html_text
        x[["loc2"]] <- doc %>%
                html_nodes(".vacborder:nth-child(4)") %>%
                html_text
        x[["loc3"]] <- doc %>%
                html_nodes(".vacborder:nth-child(5)") %>%
                html_text
        x[["epi_unit"]] <- doc %>%
                html_nodes(".vacborder:nth-child(6)") %>%
                html_text
        x[["start_date"]] <- doc %>%
                html_nodes(".vacborder:nth-child(7)") %>%
                html_text %>%
                as.Date("%d/%m/%Y")
        x[["status"]] <- doc %>%
                html_nodes(".vacborder.last") %>%
                html_text
        x[["resp"]] <- NULL
        if(!exists("P2counter")) P2counter <<- 0
        P2counter <<- P2counter + 1
        if(P2counter %% 50 == 0) {
                cat(P2counter, "\n")
                } else {
                        if(P2counter %% 10 == 0) cat("|") else cat(".")
                }
        return(data_frame(x))
        }

```

```{r function-getSO, cache=TRUE, autodep=TRUE}
#TODO
getSO <- function(SINF) {
        if(!"summary of immediate notifications and followups" %in%
                   class(SINF))
                stop("The argument \"diseaseSummaries\" is not of class \"summary of immediate notifications and followups\".")
        if(!exists("web_not_changed"))
                web_not_changed <- checkIfwebNotChanged()
#         if(!web_not_changed)
#                 warning("The OIE WAHID website has changed, the following may not work.",
#                         immediate. = TRUE)
        message("Downloading summaries of outbreaks (", nrow(SINF), "):")
        D2counter <<- 0
        summaries <- apply(SINF, 1, downloadSO)
        message("Parsing summaries of outbreaks (", length(summaries), "):")
        P2counter <<- 0
        summaries %<>% lapply(parseSO) %>%
                do.call("rbind", .)
#         attr(summaries, "disease") <- attr(diseaseSummaries, "disease")
#         attr(summaries, "countries") <- attr(diseaseSummaries, "countries")
#         attr(summaries, "years") <- attr(diseaseSummaries, "years")
        class(summaries) %<>% c(., "summary of outbreaks")
        return(summaries)
}

```

```{r getSO, cache=TRUE, autodep=TRUE}
summaries_of_outbreaks <- getSO(summaries)
write.csv2(summaries_of_outbreaks, file = "summaries_of_outbreaks.csv",
            row.names = FALSE,
            na = "",
            fileEncoding = "UTF-8")
```

```{r function-downloadOutbreakReport, cache=TRUE, autodep=TRUE}
downloadOutbreakReport <- function(x) {
        if(!all(c("reportid", "summary_country") %in% names(summaryRecord)))
                stop("Not all required arguments specified.")
        if(any(missing(year),
               missing(outbreakReportID),
               missing(outbreakCountry),
               missing(fullReportID)))
                stop("Not all required arguments specified.")
        url <- "http://www.oie.int/wahis_2/public/wahid.php/Diseaseinformation/Immsummary/outbreakreport"
        if(!exists("web_not_changed"))
                web_not_changed <- checkIfwebNotChanged()
        if(!web_not_changed)
                warning("The OIE WAHID website has changed, the following may not work.",
                        immediate. = TRUE)
        resp <- POST(url = url,
                     body = paste0("reportid=",
                                   outbreakReportID,
                                   "&summary_country=",
                                   outbreakCountry,
                                   "&backreportid=",
                                   fullReportID
                                   ),
                     content_type("application/x-www-form-urlencoded"),
                     add_headers(Referer = "http://www.oie.int/wahis_2/public/wahid.php/Diseaseinformation/Immsummary/listoutbreak")
                     )
        stop_for_status(resp)
        if(!exists("D3counter")) D3counter <<- 0
        D3counter <<- D3counter + 1
        if(D3counter %% 50 == 0) {
                cat(D3counter, "\n")
                } else {
                        if(D3counter %% 10 == 0) cat("|") else cat(".")
                }
        return(list(year = year,
                    reportid = outbreakReportID,
                    summary_country = outbreakCountry,
                    resp = resp))
        }
```

```{r function-parseOutbreakReport, cache=TRUE, autodep=TRUE}
parseOutbreakReport <- function(nonParsedOutbreakReport) {
        if(!all(c("year",
                  "summary_country",
                  "reportid",
                  "resp") %in% names(nonParsedOutbreakReport)) |
                   !"response" %in% class(nonParsedOutbreakReport[["resp"]]))
                stop ("The argument specified is not a list containing a response object.")
        url <- "http://www.oie.int/wahis_2/public/wahid.php/Diseaseinformation/Immsummary/outbreakreport"
        postfields <- rawToChar(nonParsedOutbreakReport[["resp"]]
                                [["request"]][["body"]][["config"]][["postfields"]])
        if(!all(nonParsedOutbreakReport[["resp"]][["url"]] == url,       
                grep("reportid=", postfields),
                grep("summary_country=", postfields),
                grep("backreportid=", postfields)))
                stop (paste0("The argument specified is not a non parsed outbreak report ", url))
        year <- nonParsedOutbreakReport[["year"]]
        outbreak_country <- nonParsedOutbreakReport[["summary_country"]]
        reportid <- nonParsedOutbreakReport[["reportid"]]
        doc <- html(nonParsedOutbreakReport[["resp"]], encoding = "UTF-8")
        data <- doc %>%
                html_nodes(".vacborder") %>%
                html_text
        start_date <- data[1] %>% as.Date("%d/%m/%Y")
        status <- data[2]
        resolution_date <- data[3] %>% as.Date("%d/%m/%Y")
        loc1 <- data[4]
        loc2 <- data[5]
        loc3 <- data[6]
        loc4 <- data[8]
        epi_unit <- data[7]
        lat <- data[9] %>% as.numeric
        lon <- data[10] %>% as.numeric
        description <- data[11]
        at_risk_total <- data[length(data) - 3] %>% as.numeric
        cases_total <- data[length(data) - 2] %>% as.numeric
        deaths_total <- data[length(data) - 1] %>% as.numeric
        destroyed_total <- data[length(data)] %>% as.numeric
        multi_species <- length(data) > 20
        species <- paste0(data[seq(12, length(data) - 4, 5)], collapse = ", ") # ERROR HERE
        species_breakdown <- data[12:(length(data) - 4)] %>%
                matrix(ncol = 5,
                       byrow = TRUE,
                       dimnames = list(NULL,
                                       c("species",
                                         "at_risk",
                                         "cases",
                                         "deaths",
                                         "destroyed"))) %>%
                data_frame
        species_breakdown[, 2:5] <- as.numeric(
                as.character(unlist(species_breakdown[, 2:5])))
        ret <- data_frame(year,
                          reportid,
                          outbreak_country,
                          start_date,
                          status,
                          resolution_date,
                          loc1,
                          loc2,
                          loc3,
                          loc4,
                          epi_unit,
                          lat,
                          lon,
                          description,
                          at_risk_total,
                          cases_total,
                          deaths_total,
                          destroyed_total,
                          multi_species,
                          species,
                          species_breakdown)
        if(!exists("P3counter")) P3counter <<- 0
        P3counter <<- P3counter + 1
        if(P3counter %% 50 == 0) {
                cat(P3counter, "\n")
                } else {
                        if(P3counter %% 10 == 0) cat("|") else cat(".")
                }
        return(ret)        
        }
```

```{r function-getOutbreakReports, cache=TRUE, autodep=TRUE}
getOutbreakReports <- function(outbreakSummaries, writeCSV = FALSE) {
        if(!"summary of outbreaks" %in% class(outbreakSummaries))
                stop("The argument \"outbreakSummaries\" is not of class \"summary of outbreaks\".")
        if(!exists("web_not_changed"))
                web_not_changed <- checkIfwebNotChanged()
        if(!web_not_changed)
                warning("The OIE WAHID website has changed, the following may not work.",
                        immediate. = TRUE)
        new_cache <- !file.exists("wahid_cache.sqlite3")
        wahid_cache <- src_sqlite("wahid_cache.sqlite3", create = new_cache)
        if(!new_cache) {
                cached_reports <- tbl(wahid_cache, "reports")
        }
        message(paste0("Downloading outbreak reports (",
                       nrow(outbreakSummaries),
                       "):"))
        reports <- apply(outbreakSummaries,
                         1,
                         function(x) downloadOutbreakReport(x[["year"]],
                                                            x[["outbreak_report"]],
                                                            x[["outbreak_country"]],
                                                            gsub("http://www.oie.int/wahis_2/public/wahid.php/Reviewreport/Review\\?page_refer=MapFullEventReport&reportid=",
                                                                 "",
                                                                 x[["full_report_link"]])))
        message(paste0("Parsing outbreak reports(", length(reports), "):"))
        reports %<>%
                lapply(parseOutbreakReport) %>%
                do.call("rbind", .)
        attr(reports, "disease") <- attr(outbreakSummaries, "disease")
        attr(reports, "countries") <- attr(outbreakSummaries, "countries")
        attr(reports, "years") <- attr(outbreakSummaries, "years")
        class(reports) %<>% c(., "outbreak reports")
        if(writeCSV) {
                write.table(reports,
                          file = paste0("outbreaks of ",
                                        disease,
                                        ", year ",
                                        paste0(unique(range(years)), collapse = "-"),
                                        ", countries",
                                        paste0(countries, collapse = ", "),
                                        ".txt"),
                          na = "",
                          fileEncoding = "UTF-8")
                return(NULL)
                } else {
                        return(reports)
                        }
        }
```

```{r getOutbreakReports, cache=TRUE, autodep=TRUE}
outbreak_reports <- getOutbreakReports(summaries_of_outbreaks)
write.csv2(outbreak_reports, file = "outbreak_reports.csv", row.names = FALSE,
           na = "", fileEncoding = "UTF-8")
```

```{r examplePlot, cache=TRUE, autodep=TRUE}
# library(plotGoogleMaps)
# outbreak_reportsSPDF <- outbreak_reports
# coordinates(outbreak_reportsSPDF) <- ~lon + lat
# proj4string(outbreak_reportsSPDF) <- CRS("+proj=longlat +datum=WGS84")
# ic <- iconlabels(outbreak_reportsSPDF$species, height = 5, icon = TRUE)
# map <- plotGoogleMaps(outbreak_reportsSPDF,
#                       file = "map.html",
#                       zcol = "start_date",
#                       mapTypeId = "TERRAIN",
#                       legend = FALSE,
#                       control = FALSE,
#                       iconMarker = ic)

# library(ggmap)
# theme_set(theme_bw(16))
# europe_map <- qmap("Ukraine", zoom = 5, source = "stamen")
# europe_map +
#         geom_point(aes(x = lon, y = lat, colour = start_date, size = cases_total),
#                    data = outbreak_reports)
# qmplot(lon, lat, data = outbreak_reports)

```

Download and process event summaries or full reports, if needed (contain extra information).
Cite packages!
Make a consistent flow!
Plots (animation).
Save data to a database. (SQLite)
rworldmap::countryRegions
spacetime
